---
title: "Hands-on Exercise 1.1 Geospatial Data Wrangling with R"
author: "Tan Sze Min Kara"
output: html_document
---

# 1 Overview

# 2 Getting Started

The code chunk below is used to install and load sf and tidyverse packages into the R environment.
```{r}
pacman::p_load(sf, tidyverse)
```

# 3 Importing the Data {.tabset}
## Polygon Data in Shapefile Format
```{r}
mpsz = st_read(dsn = "chap01/data/geospatial", 
                  layer = "MP14_SUBZONE_WEB_PL")
```
## Polyline Data in Shapefile Form
```{r}
cyclingpath = st_read(dsn = "chap01/data/geospatial", 
                         layer = "CyclingPathGazette")
```
## GIS data in KML format
```{r}
preschool = st_read("chap01/data/geospatial/PreSchoolsLocation.kml")
```
# 4 Checking the Content of A Simple Feature Data Frame {.tabset}

## st_geometry()
```{r}
st_geometry(mpsz)
```

## glimpse()
```{r}
glimpse(mpsz)
```

## head()
```{r}
head(mpsz, n=5)  
```

# 5 Plotting the Geospatial Data
## Plotting all features
```{r}
plot(mpsz, max.plot = 15)
```

## Plotting Only the Geometric Shapes
```{r}
plot(st_geometry(mpsz))
```

## Plotting using a specific attribute
```{r}
plot(mpsz["PLN_AREA_N"])
```

# 6 Working with Projection {.tabset}
## Assigning EPSG code to a simple feature data frame
### Default Code
```{r}
st_crs(mpsz)
```

### Corrected Code
```{r}
mpsz3414 <- st_set_crs(mpsz, 3414)

st_crs(mpsz3414)
```

## Transforming the projection of preschool from wgs84 to svy21.
```{r}
preschool3414 <- st_transform(preschool, 
                              crs = 3414)
st_geometry(preschool3414)
```

# 7 Importing and Converting An Aspatial Data
## Importing the aspatial data

```{r}
listings <- read_csv("chap01/data/aspatial/listings.csv")

list(listings) 
```

## Creating a simple feature data frame from an aspatial data frame
```{r}
listings_sf <- st_as_sf(listings, 
                       coords = c("longitude", "latitude"),
                       crs=4326) %>%
  st_transform(crs = 3414)

glimpse(listings_sf)
```

# 8 Geoprocessing with sf package
## Buffering
```{r}
buffer_cycling <- st_buffer(cyclingpath, 
                               dist=5, nQuadSegs = 30)

buffer_cycling$AREA <- st_area(buffer_cycling)

sum(buffer_cycling$AREA)
```

## Point-in-polygon count
```{r}
mpsz3414$`PreSch Count`<- lengths(st_intersects(mpsz3414, preschool3414))

summary(mpsz3414$`PreSch Count`)

top_n(mpsz3414, 1, `PreSch Count`)
```

# 9  Exploratory Data Analysis (EDA)
## Histogram hist()
```{r}
mpsz3414$Area <- mpsz3414 %>%   st_area()

mpsz3414 <- mpsz3414 %>%
  mutate(`PreSch Density` = `PreSch Count`/Area * 1000000)

hist(mpsz3414$`PreSch Density`)
```

## Histogram ggplot()
```{r}
ggplot(data=mpsz3414, 
       aes(x= as.numeric(`PreSch Density`)))+
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue") +
  labs(title = "Are pre-school even distributed in Singapore?",
       subtitle= "There are many planning sub-zones with a single pre-school, on the other hand, \nthere are two planning sub-zones with at least 20 pre-schools",
      x = "Pre-school density (per km sq)",
      y = "Frequency")
```

### Scatterplot ggplot()
```{r}
ggplot(data=mpsz3414, 
       aes(y = `PreSch Count`, 
           x= as.numeric(`PreSch Density`)))+
  geom_point(color="black", 
             fill="light blue") +
  xlim(0, 40) +
  ylim(0, 40) +
  labs(title = "",
      x = "Pre-school density (per km sq)",
      y = "Pre-school count")
```
